MLH1 Results Outline

### 2020-02-18 doc/Results.rmd
small changes to ordering and geting numbers to appear

In order to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` _M. m. musculus_ strains, MLH1 foci were quantified from a total of `r sum( (main.table  %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table  %>% filter(sex == "female"))$Ncells)` oocytes.

(mouse means were calculated and statisics in Table 2)

Figure 1(A) shows the --

--notes on error /

-- connections to previous measures in the literature

<!--
The strain means for females were pretty similar. For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively. In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively. In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively. 
-->

@@

# Female Specific Analysis

In order to more finely describe the sex specific patterns, we focused on data sets including single sexes.

female.specific.plot

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
F.strain.table

For some strains, we were only able to quantified or include one mouse and the Cast observation comes from 1 1 cell.
To analyze the female specific data set,

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _domesticus^G^_

In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _ musculus^PWD^_,`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_

@@

#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co)
```

@@

_domesticus^G^_ is `r mean.G.female / mean.all.other` higher the the other means. _ domesticu^LEW^_  is `r mean.LEW.female / mean.all.other` higher  and _musculus^MSM^_ is `r mean.MSM.female / mean.all.other`. These three will be designated as 'moderate high rec' strains for later sex specific analysis.

@@

# Male Specific Analysis

The male specific patterns were also investigated. (what's a better motivation for this section)

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
#try customize the plot size

```{r male.strain.table, echo=FALSE}

M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

                        #Standard error (mice . cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
M.strain.table
#same issue as the female table .. ugh
```

The same model framework was applied to the male specific data.
The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.

 _domesticus^LEW^_

@@

The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively. Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.

>>>>

In order to estimate the genome wide recombination rates across our panel of three subspecies from the house mouse species complex, _M. m. domesticus,_ _M. m. musculus_. and _M. m. molossinus_. 
`r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains were used with reflect samples of natural genetic variation across the terretories of these subspecies.

MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes. The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises.

1. -female strain means are v close  -male strain means are more varied!!
-Comparisons to references / literatures
-house mouse close to minimum of 1  obligate CO per chrm

2. The largest female difference is G, 1.07X  the largest male difference is PWD and MSM which are 1.3 and 1.2 respectively.

<!-- female -->
In order to more finely describe the sex specific patterns, we focused on data sets including single sexes.

#female.specific.plot

## female section

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#F.strain.table

```{r female.dif, echo=FALSE}
#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co)

G.female.diff <- (mean.G.female / mean.all.other)
LEW.female.diff <- (mean.LEW.female / mean.all.other)
MSM.female.diff <- (mean.MSM.female / mean.all.other)

The overall pattern for the female mean MLH1 counts per cell is low variation. ( in the most different strains
In females specifically _domesticus^G^_ is `r G.female.diff` higher the the other means. 
while _ domesticu^LEW^_  is `r LEW.female.diff` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )`

in _domesticus^G^_

In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _musculus^PWD^_,

`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_

### male section

While the male pattern shows more variation across the strain means.

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
#try customize the plot size

```{r male.strain.table, echo=FALSE}
M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

                        #Standard error (mice . cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#M.strain.table
#same issue as the female table .. ugh
```

```{r male.diff, echo=FALSE}

The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r (mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co) )`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively. Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.

--notes on error / 
-- connections to previous measures in the literature

The same model framework was applied to the male specific data.
The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
<!--
The strain means for females were pretty similar. For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively. In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively. In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively. 
-->

@@

MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS

@@

@@

@@

MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS

@@
